---
title: "Web Scraping and Data Cleaning"
output: github_document
---

```{r setup, include=FALSE}
library(rvest)
library(stringr)
library(tidyverse)
```

## Project

This project uses the rvest package to retrieve the data stored in an HTML table at,
https://oldschool.runescape.com/slu. This webpage contains data about the game's servers and the number of players on them. This data will be cleaned and used in some quick visualizations. 


## Web Scrape

The rvest package is used to retreive the data stored in an HTML table. 

```{r}
rsweb <- read_html("https://oldschool.runescape.com/slu")

rswebdata <- html_nodes(rsweb,'.server-list__row-cell')
rsdata <- html_text(rswebdata)
rsdata <- gsub("\n"," ",rsdata)
head(rsdata)
```

## Data Cleaning

Due to the coding of the HTML on the Runescape website, the data is scraped into list format. Each row contains 5 variables so the split function is used to create a row for every five values in the list.

```{r}
rsdata <- as.data.frame(split(rsdata, 1:5))
head(rsdata, n=15)

```

The default column names are very unhelpful, so they are changed to more descriptive names.

```{r}
rsdata <- rsdata %>%
  rename(world = X1, players = X2, country = X3, type = X4, activity = X5)

colnames(rsdata)
```
Most of the values contain some unneeded characters. "OldSchool," is not needed for the world names coloumn, some rows also have spaces and others do not. The gsub function is used to remove spaces and "OldSchool," leaving just the number. In the players coloumn, "players," is not needed and can be removed so it can be later converted to an integer datatype. Finally, if a world has no specific activity it only has a "-" this is changed to "regular," to be more clear. 

```{r}
rsdata$world <- gsub(" ","",rsdata$world)
rsdata$world <- gsub("OldSchool","",rsdata$world)

rsdata$players <- gsub(" ","",rsdata$players)
rsdata$players <- gsub("players","",rsdata$players)

rsdata$activity <- gsub("-","regular",rsdata$activity)

rsdata$players <- as.integer(rsdata$players)


str(rsdata)
head(rsdata)

```
## visualizations

Looking at country, most of the servers are for the United states, including by far the highest number of premium, "member," servers. 

```{r}
ggplot(rsdata, aes(x=type, fill=country)) + geom_bar(position = "dodge")

```

Accordingly, the United States also has the highest number of players, both free and member. The United States has a much higher number of members in their playerbase compared to other countries.
```{r}
playercount <- rsdata %>%
  group_by(country, type) %>%
  summarize(totalplayers=sum(players))

ggplot(playercount, aes(x=country, y=totalplayers, col=type)) + geom_point()

```

The differential between the free and member countries is fairly similar between all non-United States countries. The United States has by far the highest portion of their player base paying to be a member, making the United States the most profitable player base as well.
```{r}
playercountspread <- spread(playercount, type, totalplayers)

playercountspread$playertypediff <- playercountspread$Members - playercountspread$Free
playercountspread

ggplot(playercountspread, aes(x=reorder(country, +playertypediff), y=playertypediff, fill=country)) + geom_col(position = "dodge") +xlab("Countries")+ ylab("Player Type Differential")
```
